The Heavy Photon Search Test Detector

### 2014-03-25 hps-testrun-nim.tex
Changes from large overhaul by John and myself.

 Data was taken on three different converter thicknesses with a beam current varying between 
30-70~nA and repeated with the reverse field setting of the PS dipole magnet.
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on converter \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the test run produced a relatively large fraction of pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test is summarized in Tab.~\ref{tab:currents}.

@@

The design of the SVT is primarily driven by direct physics requirement and constraints from the 
environment at the interaction region. The \Aprime decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime's is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10 cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5 mm from the center of the beam, 
defining the 15 mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$ electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, as shown in 
Fig.~\ref{fig:radiation},  requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
\begin{figure}[]
\includegraphics[width=7cm]{figures/radiation}
\label{fig:radiation}
\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

@@

%\subsubsection{Front-end Electronics}
Despite having only small spots with very high occupancy (up to 4~MHz/mm$^2$) closest to the primary 
beam, see Fig.~\ref{fig:radiation}, the rates are still high and lowering the peak occupancy to 
approximately 1\% for tracking requires a trigger window and hit time tagging of roughly 8~ns. The 
ECal readout and trigger described in Sec.~\ref{sec:fadc} can achieve such resolution. To reach this 
performance the sensors for the SVT are readout by the APV25 ASIC developed for the CMS 
experiment at CERN~\cite{French:2001xb}. The APV25 can capture two successive samples of three 
of the output of the shaper at a sampling rate of approximately 40~MHz.  By fitting the known 
$CR$-$RC$ shaping curve to these samples, the initial time of the hit can be determined to a precision 
of 2~ns for S/N$\approx25$~\cite{Friedl:2009zz}.  For electron beam running, six-sample readout and 
the shortest possible shaping time (35~ns) is used to best distinguish hits that overlap in time.
The APV25 ASICs are hosted on simple FR4 hybrid readout boards, outside the tracking volume, with a 
short twisted-pair pigtail cable to provide power and configuration and signal readout. Along with a 
single sensor, these are glued to a polyamide-laminated carbon fiber composite backing making 
up a half-module. A window is machined in the carbon fiber leaving only a frame around the periphery 
of the silicon to minimize material. A 50~$\mu$m sheet of polyamide is laminated to the surface of the 
carbon fiber with 1~mm overhang at all openings to ensure good isolation between the backside of the 
sensor, carrying high-voltage bias, and the carbon fiber which is held near ground.

@@

During the Test run, both halves were held in place by four vertical bars attached to an above rail, 
placing the front face of the crystals 147~cm from the upstream edge of the magnet and with a 
8.7~cm gap between the innermost edge of the crystals in the two halves.
{\small
\begin{figure*}[htb]
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/ECal}
\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
\caption{Rendered layout view (left) and the ECal assembly on the beam line (right).}
\label{fig:ecal}
\end{center}
\end{figure*}
}

\subsection{Signal readout}
\label{sec:fadc}
After a 2:1 signal splitter, $1/3$ of an amplified APD signal ($2/3$ is sent to a timing module) is fed to a 
single channel of a JLab flash ADC (FADC) board~\cite{fadc}, 
see Fig~.\ref{fig:fadc}.
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=4cm]{figures/FADC250_Photo_001.jpg}
\caption{The JLab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
The FADC boards are high speed VXS modules digitizing up to 16 APD signals 
at 250~MHz and storing samples in 8~$\mu$s deep pipelines with 12-bit resolution. 
When a trigger is received, the part 
of the pipeline from 5 samples before and 30 after the signal crossed a programmable 
threshold (for the Test run this was set to $\approx70$~MeV) are 
summed and stored in a 17-bit register for readout as described in Fig.~\ref{fig:hps_trigger_data}.
This scheme significantly compresses the data output of the FADC. During offline data analysis, a 
calibrated pedestal value is subtracted to obtain the actual summed energy.
Two 20-slot VXS crates with 14 (13) FADC boards were employed in the Test run to read out the top 
(bottom) half of the ECal.
In the Test run 385 out of 442 modules (87\%) was used in offline reconstruction, 39 modules were 
disabled or not read out (no FADC channel available, no APD bias voltage or masked out due to 
excessive noise) and 18 was masked offline due to noise. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=8cm]{figures/fadc_datapath_diagram.png}
\caption{The two different readout modes of the FADC used in the Test run.}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

@@

\section{Trigger and Data Acquisition}

@@

@@

After a trigger is received, six samples of the corresponding output of the APV25 ASIC shaper circuit are 
digitized. The six pedestal-subtracted samples from every channel in the SVT are the basis for offline hit 
reconstruction. The typical pulse shape obtained shown in Fig.~\ref{fig:pulse_shape} also 
demonstrates that the SVT was well timed in to the trigger with the rise of the pulse at the 3rd sampling 
point.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{Accumulation of six pedestal-subtracted samples from individual SVT channels associated 
	with hits on tracks.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}
These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips with the position of a cluster on the sensor determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[]
\begin{center}
{\small
	%\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
	\includegraphics[width=7cm]{figures/mip_top_layer_2.png}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

@@

By selecting \ee{} pairs from the triggered events we are able to study basic distributions of pair 
production kinematics. Pairs of opposite charge tracks, one in the top and one in the bottom half of the 
SVT, with momentum larger than 400~MeV were selected. The pair production kinematics are relatively 
well reproduced given the alignment of the tracker as shown in Fig.~\ref{fig:pair_kin}. 
 \begin{center}
{\small
\begin{figure*}[t]
   \includegraphics[width=5cm]{figures/h_trk_top_px_h_trk_top_px_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_invM_h_invM_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_sumE_h_sumE_trigsel4hit_pair1351_twotrkfilt-v6-paper}
\caption{Kinematic distributions for e$^+$e$^-$ pairs selected by opposite charged tracks in the top and bottom half of the tracker: track momentum in the top half of the SVT (left), invariant mass (middle) and the sum of the track momentum for the pair (right).}
\label{fig:pair_kin}
\end{figure*}
}
\end{center}
The expected momentum resolution from simulation is between 4-5\% 
for tracks in the momentum range of the Test run. This is significantly smaller than the accuracy at 
which we're able estimate the scale and resolution in the Test run. By looking at the agreement 
between data and simulation in the shape of the kinematic distributions for single- and two track 
events and adjusting the overall momentum scale and smearing simulated events with a 
Gaussian resolution we estimate an uncertainty of 10\% on both the momentum scale and 
resolution.

For the vertexing performance the foremost difference compared to electron beam running is that the 
target was located approximately 67~cm upstream from our nominal target position; giving almost 
collinear tracks in the detector. This degrades the vertex resolution along the 
beam line compared to that expected in an electron beam with nominal target position. 
Furthermore, tails of the vertex distributions are impossible to study with the finite data sample obtained 
in the Test run. Nevertheless, useful information can still be 
obtained by studying the vertex distributions. Figure~\ref{fig:vtx_pos} shows the distance of closest 
approach of the momentum vectors extrapolated in the upstream direction from our analyzing magnet, 
taking into account the measured fringe field of the PS dipole magnet. 
 \begin{center}
{\small
	 \begin{figure*}[t]
	% Plot info:
	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
	\includegraphics[width=5cm]{figures/h_vtx_fr_x_h_vtx_x_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\includegraphics[width=5cm]{figures/h_vtx_fr_y_h_vtx_y_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\includegraphics[width=5cm]{figures/h_vtx_fr_z_h_vtx_z_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\caption{ Vertex position represented by the distance of closest approach of the extrapolated 
	momentum vectors upstream of the PS dipole magnet taking into account the measured fringe field. 
	The overall shift from zero in the x-direction is due to a 30~mrad rotation of the SVT with respect to 
	the beam line.}
	\label{fig:vtx_pos}
\end{figure*}
}
\end{center}

@@

\subsection{Multiple Coulomb Scattering Distributions}
\label{sec:mcs}
One of the main goals of the Test run was to evaluate the description of the tails of the 
multiple Coulomb scattering to gain further confidence in the expected detector occupancy in the full HPS 
experiment. Those occupancies have been evaluated using \egs{}~\cite{egs5} simulation program 
whose \moliere{} formulation of multiple scattering show good agreement with previous studies 
for a wide range of materials and projectiles~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467}. 
Despite all data taken with a photon beam in the Test run, the angular distribution of the pair 
produced electron and positron emerging from the converter with different converter thickness can 
be used to study the varying multiple Coulomb scattering contribution (pair production angle is constant). 
Restricting clusters to a well calibrated region of the ECal we observe, within our systematic 
uncertainties , good agreement with the rates predicted by \egs{}, see Tab.~\ref{tab:mcs}.
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
GEANT4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted number of events for 1~s of beam at 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.}
}
\label{tab:mcs}
\end{table}
%\end{center}

\subsection{Trigger Performance}
As described in Sec.~\ref{sec:fadc} and shown in Fig.~\ref{fig:hps_trigger_data} the energy from each 
crystals are measured differently in the trigger and what is readout from the ECal. 
The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC trigger path by converting from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision from the SSP is simulated before we compare the trigger decision 
and trigger time to what was reported by the actual trigger. For every event, the trigger reports the trigger 
decision as a bit mask (top half, bottom half or both) and the time the trigger fired. To eliminate trigger 
bias, we use a tag and probe method: to study the trigger performance in one half of the ECal, we select 
events which triggered the other half and where there was exactly one probe cluster in the ECal half 
under study. We then measure trigger efficiency as the fraction of tagged events that fired the trigger in 
the probe half as a function of the probe cluster energy, shown in Fig.~\ref{fig:turnon}. 
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/h_cl_E_probedata_eff_h_cl_E_probeMC_eff_dataMC_1351-v7-trig-tagBot.pdf}
	\caption{Trigger efficiency in both halves of the ECal for data and simulation as a 
	function of cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below 1~GeV for two reasons;  
gain variations between different crystals lead to the threshold variations and the nonlinearity of 
the time-over-threshold integral means that the effective threshold is higher for clusters that span 
multiple crystals. The effective trigger threshold is therefore dependent on position and energy of 
the particle as well as cluster multiplicity.

@@

@@

\section{Acknowledgements}

>>>>

 Data was taken on three different converter thicknesses with a beam current varying between 
30-70~nA and repeated with the reverse field setting of the PS dipole magnet.
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the test run produced a relatively large fraction of pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test is summarized in Tab.~\ref{tab:currents}.

@@

The design of the SVT is primarily driven by direct physics requirement and constraints from the 
environment at the interaction region. The \Aprime decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime's is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10 cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5 mm from the center of the beam, 
defining the 15 mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$ electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

@@

%\subsubsection{Front-end Electronics}
Despite having only small spots with very high occupancy (up to 4~MHz/mm$^2$) closest to the primary 
beam, the rates are still high and lowering the peak occupancy to 
approximately 1\% for tracking requires a trigger window and hit time tagging of roughly 8~ns. The 
ECal readout and trigger described in Sec.~\ref{sec:fadc} can achieve such resolution. To reach this 
performance the sensors for the SVT are readout by the APV25 ASIC developed for the CMS 
experiment at CERN~\cite{French:2001xb}. The APV25 can capture two successive samples of three 
of the output of the shaper at a sampling rate of approximately 40~MHz.  By fitting the known 
$CR$-$RC$ shaping curve to these samples, the initial time of the hit can be determined to a precision 
of 2~ns for S/N$\approx25$~\cite{Friedl:2009zz}.  For electron beam running, six-sample readout and 
the shortest possible shaping time (35~ns) is used to best distinguish hits that overlap in time.
The APV25 ASICs are hosted on simple FR4 hybrid readout boards, outside the tracking volume, with a 
short twisted-pair pigtail cable to provide power and configuration and signal readout. Along with a 
single sensor, these are glued to a polyamide-laminated carbon fiber composite backing making 
up a half-module. A window is machined in the carbon fiber leaving only a frame around the periphery 
of the silicon to minimize material. A 50~$\mu$m sheet of polyamide is laminated to the surface of the 
carbon fiber with 1~mm overhang at all openings to ensure good isolation between the backside of the 
sensor, carrying high-voltage bias, and the carbon fiber which is held near ground.

@@

During the Test run, both halves were held in place by four vertical bars attached to an above rail, 
placing the front face of the crystals 147~cm from the upstream edge of the magnet and with a 
8.7~cm gap between the innermost edge of the crystals in the two halves.
{\small
\begin{figure}[]
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/ECal}
%\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
\caption{Rendered layout view of the ECal looking downstream.}
\label{fig:ecal}
\end{center}
\end{figure}
}

\subsection{Signal readout}
\label{sec:fadc}
After a 2:1 signal splitter, $1/3$ of an amplified APD signal ($2/3$ is sent to a timing module) is fed to a 
single channel of a JLab flash ADC (FADC) board~\cite{fadc}.
%\begin{figure}[]
%\begin{center}
%{\small
%\includegraphics[width=4cm]{figures/FADC250_Photo_001.jpg}
%\caption{The JLab FADC250 VXS module.}
%\label{fig:fadc}
%}
%\end{center}
%\end{figure}
The FADC boards are high speed VXS modules digitizing up to 16 APD signals 
at 250~MHz and storing samples in 8~$\mu$s deep pipelines with 12-bit resolution. 
When a trigger is received, the part 
of the pipeline from 5 samples before and 30 after the signal crossed a programmable 
threshold (for the Test run this was set to $\approx70$~MeV) are 
summed and stored in a 17-bit register for readout.
This scheme significantly compresses the data output of the FADC. During offline data analysis, a 
calibrated pedestal value is subtracted to obtain the actual summed energy.
Two 20-slot VXS crates with 14 (13) FADC boards were employed in the Test run to read out the top 
(bottom) half of the ECal.
In the Test run 385 out of 442 modules (87\%) was used in offline reconstruction, 39 modules were 
disabled or not read out (no FADC channel available, no APD bias voltage or masked out due to 
excessive noise) and 18 was masked offline due to noise. 
%\begin{figure}[]
%\begin{center}
%{\small
%\includegraphics[width=8cm]{figures/fadc_datapath_diagram.png}
%\caption{The two different readout modes of the FADC used in the Test run.}
%\label{fig:hps_trigger_data}
%}
%\end{center}
%\end{figure}

@@

\section{Trigger and Data Acquisition}
\label{sec:triggerdaq}

@@

In order to minimize cable length for the analog APV25 output signal the ATCA crate was located 
approximately 1~m from the beam line, next to our cable vacuum feed troughs.   
Before shielding with lead-blankets was arranged, we observed two failures of normally reliable ATCA 
crate power supplies, time-correlated to beam instabilities.

@@

After a trigger is received, six samples of the corresponding output of the APV25 ASIC shaper circuit are 
digitized. The six pedestal-subtracted samples from every channel in the SVT are the basis for offline hit 
reconstruction. The typical pulse shape obtained shown in Fig.~\ref{fig:pulse_shape} also 
demonstrates that the SVT was well timed in to the trigger with the rise of the pulse at the 3rd sampling 
point.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{Accumulation of six pedestal-subtracted samples from individual SVT channels associated 
	with hits on tracks.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}
These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips with the position of a cluster on the sensor determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape. 
%, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
%\begin{figure}[]
%\begin{center}
%{\small
%	%\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
%	\includegraphics[width=7cm]{figures/mip_top_layer_2.png}
%    	\caption{ The cluster charge distribution for hits associated with a track. }
%	\label{fig:cluster_pulse}
%}
%\end{center}
%\end{figure}

@@

By selecting \ee{} pairs from the triggered events we are able to study basic distributions of pair 
production kinematics. Pairs of opposite charge tracks, one in the top and one in the bottom half of the 
SVT, with momentum larger than 400~MeV were selected. The pair production kinematics are relatively 
well reproduced given the alignment of the tracker as shown in Fig.~\ref{fig:pair_kin}.

The expected momentum resolution from simulation is between 4-5\% 
for tracks in the momentum range of the Test run. This is significantly smaller than the accuracy at 
which we're able estimate the scale and resolution in the Test run. By looking at the agreement 
between data and simulation in the shape of the kinematic distributions for single- and two track 
events and adjusting the overall momentum scale and smearing simulated events with a 
Gaussian resolution we estimate an uncertainty of 10\% on both the momentum scale and 
resolution.
 \begin{center}
{\small
\begin{figure*}[t]
   \includegraphics[width=5cm]{figures/h_trk_top_px_h_trk_top_px_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_invM_h_invM_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_sumE_h_sumE_trigsel4hit_pair1351_twotrkfilt-v6-paper}
\caption{Kinematic distributions for e$^+$e$^-$ pairs selected by opposite charged tracks in the top and bottom half of the tracker: track momentum in the top half of the SVT (left), invariant mass (middle) and the sum of the track momentum for the pair (right).}
\label{fig:pair_kin}
\end{figure*}
}
\end{center}
In the Test run, as well as in electron running with HPS, the dominating source of uncertainty in the 
tracking and vertexing is multiple Coulomb scattering. For the vertexing performance the foremost 
difference compared to electron beam running is that the target was located approximately 67~cm 
upstream from our nominal target position; giving almost collinear tracks in the detector. The increased 
lever arm over which tracks are extrapolated widens the resolution with up to a factor of eight (depending 
on momentum) compared to what is achieved at the nominal electron target position for HPS. 
Figure~\ref{fig:impact_param} shows the horizontal and vertical positions of the extrapolated track at the 
converter position. 
{\small
	 \begin{figure}[]
 \begin{center}
	% Plot info:
	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
	\includegraphics[width=7cm]{figures/h_trk_top_fr_conv_z_h_trk_top_conv_z_dataMC_twotrksel.png}
	\includegraphics[width=7cm]{figures/h_trk_top_fr_conv_y_h_trk_top_conv_y_dataMC_twotrksel.png}
	\caption{ Vertical (top) and horizontal (bottom) extrapolated track position at the converter position 
	taking into account the measured fringe field. }
	\label{fig:impact_param}
\end{center}
\end{figure}
}
While residual alignment show small shifts good agreement on the widths between data and simulated 
events, indicating a good understanding of our material budget and distribution in the SVT.  
%Furthermore, tails of the vertex distributions are impossible to study with the finite data sample obtained 
%in the Test run. Nevertheless, useful information can still be 
%obtained by studying the vertex distributions. Figure~\ref{fig:vtx_pos} shows the distance of closest 
%approach of the momentum vectors extrapolated in the upstream direction from our analyzing magnet, 
%taking into account the measured fringe field of the PS dipole magnet. 
 %\begin{center}
%{\small
%	 \begin{figure*}[t]
%	% Plot info:
%	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
%	\includegraphics[width=5cm]{figures/h_vtx_fr_x_h_vtx_x_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\includegraphics[width=5cm]{figures/h_vtx_fr_y_h_vtx_y_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\includegraphics[width=5cm]{figures/h_vtx_fr_z_h_vtx_z_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\caption{ Vertex position represented by the distance of closest approach of the extrapolated 
%	momentum vectors upstream of the PS dipole magnet taking into account the measured fringe field. 
%	The overall shift from zero in the x-direction is due to a 30~mrad rotation of the SVT with respect to 
%	the beam line.}
%	\label{fig:vtx_pos}
%\end{figure*}
%}
%\end{center}

@@

\subsection{Trigger Performance}
As described above the energy from each 
crystal is measured differently in the trigger and what is readout from the ECal. 
The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC trigger path by converting from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision from the SSP is simulated before we compare the trigger decision 
and trigger time to what was reported by the actual trigger. For every event, the trigger reports the trigger 
decision as a bit mask (top half, bottom half or both) and the time the trigger fired. To eliminate trigger 
bias, we use a tag and probe method: to study the trigger performance in one half of the ECal, we select 
events which triggered the other half and where there was exactly one probe cluster in the ECal half 
under study. We then measure trigger efficiency as the fraction of tagged events that fired the trigger in 
the probe half as a function of the probe cluster energy, shown in Fig.~\ref{fig:turnon}. 
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/h_cl_E_probedata_eff_h_cl_E_probeMC_eff_dataMC_1351-v7-trig-tagBot.pdf}
	\caption{Trigger efficiency in both halves of the ECal for data and simulation as a 
	function of cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below 1~GeV for two reasons;  
gain variations between different crystals lead to the threshold variations and the nonlinearity of 
the time-over-threshold integral means that the effective threshold is higher for clusters that span 
multiple crystals. The effective trigger threshold is therefore dependent on position and energy of 
the particle as well as cluster multiplicity.

@@

\subsection{Trigger Rate Comparisons}
\label{sec:mcs}
Trigger rates observed in the Test run are dominated by multiple scattered pair-produced \ee pairs in the 
converter. In simulated events, the rate of triggers depend on the modeling of the pairs angular distribution 
and the sub-sequent multiple Coulomb scattering in the converter. Rates from different converter 
thicknesses are used to study the varying multiple Coulomb scattering contribution (pair production 
angle is constant). Restricting clusters to a well calibrated region of the ECal and subtracting the 
"no converter" background we see agreement with the rates predicted by the \egs{} simulation 
program, see Tab.~\ref{tab:mcs}.
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
GEANT4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted number of events for 1~s of beam at 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.}
}
\label{tab:mcs}
\end{table}
%\end{center}
This gives further confidence that the dominant source of background occupancy for HPS, multiple Coulomb 
scattered beam electrons, is well described. 
%One of the main goals of the Test run was to evaluate the description of the tails of the 
%multiple Coulomb scattering to gain further confidence in the expected detector occupancy in the full HPS 
%experiment. Those occupancies have been evaluated using \egs{}~\cite{egs5} simulation program 
%whose \moliere{} formulation of multiple scattering show good agreement with previous studies 
%for a wide range of materials and projectiles~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467}. 
%Despite all data taken with a photon beam in the Test run,

@@

\section{Acknowledgements}
The authors are grateful for the support from Hall~B at JLab and especially Hall~B technicians 
for support during installation and decommissioning. They also would like to commend the CEBAF 
personnel for achieving good performance when we needed it; the last few hours of operating CEBAF6. 
The tremendous support from home institutions and supporting staff also needs praise from the authors.

This work has been supported by the US Department of Energy.
